---
title: "Training Peaks"
output:
  html_notebook:
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: true
    theme: cosmo
    df_print: paged
    highlight: tango
    # code_folding: hide
# output:
#   epuRate::BAKER:
#     toc: TRUE
#     number_sections: FALSE
#     code_folding: "show"
---

<script>
$(document).ready(function() {
  $items = $('div#TOC li');
  $items.each(function(idx) {
    num_ul = $(this).parentsUntil('#TOC').length;
    $(this).css({'text-indent': num_ul * 10, 'padding-left': 0});
  });

});
</script>

***

# Play around with FIT data

## Load Libraries

```{r warning=FALSE, results='hide'}
library(fit)
library(ggplot2)
library(leaflet)
library(dplyr)
library(cetcolor)

```

*** 

# Tutorial

Data from https://github.com/russelldb/fit


```{r}
data <- read.fit('soft/fit-R/examples/mt_beauty.fit')
# data$record$distance <- data$record$distance/100
#weird function to calculate correct altitude compared to tutorial
# data$record$altitude <- unlist(lapply(data$record$altitude,function(x) x/(14.35-(x-3838)*6/2800)))
names(data)
head(data$record)
attr(data$record,'units')
```


```{r}
library(ggplot2)
```


```{r}
pdata <- with(data$record, data.frame(alt = altitude, time = (timestamp-timestamp[1])/60))
ggplot(pdata, aes(y=alt, x=time)) + geom_line() +
  ggtitle("Elevation vs Time") + xlab("time (minutes)") + ylab("elevation (m)")
```


```{r}
pdata <- with(data$record, data.frame(alt = altitude, time = (distance-distance[1])/1000))
ggplot(pdata, aes(y=alt, x=time)) + geom_line() +
  ggtitle("Elevation vs Distance") + xlab("Distance (km)") + ylab("elevation (m)")
```
```{r}
pdata <- data$record[-(1:10),c("heart_rate","timestamp")]
# compute average gradient, as %
pdata$gradient <- with(data$record, 100 * diff(altitude,lag=10) / diff(distance,lag=10))
pdata <- subset(pdata, complete.cases(pdata) & abs(gradient) < 7.5 & gradient != 0) # drop outliers
ggplot(pdata, aes(x=gradient, y=heart_rate)) + 
  geom_point(alpha=0.5) + geom_jitter() +
  stat_smooth(method="lm", formula=y ~ poly(x, 2)) +
  ggtitle("Heart rate vs gradient")
```

What's really interesting is that the relationship is non-linear: as the road gets steeper, the marginal increase in heart rate diminishes. This is probably because there's only so hard my heart can actually beat---and rather than suffer cardiac arrest, I slow the bike down a bit and let my heart muscle rest.

In the graph above, I've approximated the relationship with a quadratic function. Let's look closer at the mapping between heart rate and road gradient:

```{r}
fit <- lm(heart_rate ~ poly(gradient, 2), data=pdata)
summary(fit)
```
```{r}
pred <- data.frame(gradient = seq(-6,6,2))
pred$heart_rate <- predict(fit, pred)
t(pred)
```

The regression output above says that the road's gradient alone explains about 70% of variation in my average heart rate. Riding on flat road that day, my heart could be expected to be beating at about 175bpm (the model intercept). When riding uphill, my heart worked harder---a larger gradient was associated with a faster heartbeat, although the increase is diminishing for each incremental point of gradient. An increase to a 2% gradient would increase my heart rate by 4bpm to about 180, but a further 2 point increase to 4% would see just a 2½bpm increase. On the other hand, a decrease to a -2% gradient road would be associated with a larger 5bpm decline in average heart rate, to 170bpm. Coasting down the really steep bits, at -6% average gradient, my heart was pumping at just 157bpm.

## Making maps

We have GPS data, and it'd be a shame not to put the data on a map. One way to do this is to use David Kahle and Hadley Wickham's excellent ggmap package. In the map below, we get a map image from Google Maps, and overlay the path I travelled, coloring the path according to my heart rate at that section:

```{r}
# library(ggmap)
# 
# garmin.data <- read.fit('soft/fit-R/examples/mt_beauty.fit')
points <- subset(data$record, complete.cases(data$record))
points$time_min  <- with(points, timestamp - timestamp[1])/60 # minutes of riding

# from diagram above, we turned around at the 90 minutes mark
points[which(points$altitude == max(points$altitude)),]
```

```{r}
points$direction <- with(points, factor(ifelse(time_min < 90, 'Outbound', 'Return')))

# map <- get_googlemap(center = colMeans(points[,c('position_long','position_lat')]),
#                      zoom = 12, maptype = c("hybrid"))
# 
# ggmap(map, extent='panel') +
#   geom_path(aes(x = position_long, y = position_lat, colour = heart_rate),
#              data = points, size = 2) +
#   facet_grid(direction~.) +
#   scale_colour_gradientn(colours = rainbow(3)) +
#   ggtitle("Mt Beauty, 23 January 2014") + ylab('latitude') + xlab('longitude')
```

Does not work currently, as GOOGLEMAPS asks for an API (with associated costs), will try to find alternatives:

```{r}
library(cetcolor)
ggplot(points,
       aes(y=position_lat, x=position_long, color=heart_rate)) +
  geom_path() + theme_dark() +
  facet_grid(direction~.) +
  xlab('Longitude') + ylab('Latitude') +
  ggtitle('Forerunner 230 GPS and Heart Rate Data') +
  scale_color_gradientn('Heart Rate (bpm)', colours = cet_pal(5, name="inferno")) +
  theme(plot.title = element_text(hjust=0.5, size=rel(2)),
        axis.title = element_text(size=rel(2)),
        legend.title = element_text(size=rel(1.5))) +
  ylim(c(-36.80,-36.65)) 
```

Checking whether "inferno" is "l8" in linear palettes

```{r}
display_cet_attribute(attribute = "linear")
# display_cet_all()
# ?cet_color_maps
```


```{r}
ggplot(points,
       aes(y=position_lat, x=position_long, color=heart_rate)) +
  geom_path() + theme_dark() +
  facet_grid(direction~.) +
  xlab('Longitude') + ylab('Latitude') +
  ggtitle('Forerunner 230 GPS and Heart Rate Data') +
  scale_color_gradientn('Heart Rate (bpm)', colours = rev(cet_pal(5, name="l3"))) +
  theme(plot.title = element_text(hjust=0.5, size=rel(2)),
        axis.title = element_text(size=rel(2)),
        legend.title = element_text(size=rel(1.5))) +
  ylim(c(-36.80,-36.65)) 
```

## With Leaflet

How to do it with an actual map?

* From
  * https://cran.r-project.org/web/packages/cycleRtools/vignettes/cycling_data_analysis.html
* To add Leaflet gradient
  * https://stackoverflow.com/questions/50806293/add-gradient-color-to-polyline-in-leaflet-r

```{r}
library(leaflet)
leaflet(points[points$direction == 'Outbound',]) %>% addTiles() %>% addPolylines(~position_long,~position_lat)
leaflet(points[points$direction == 'Return',]) %>% addTiles() %>% addPolylines(~position_long,~position_lat)
```
```{r}
newcols <- rev(cet_pal(min(dim(points)[1],256),name="l3"))
points <- points %>% mutate(quantile=ntile(heart_rate,256))
newcols.quantile <- newcols[points$quantile]
points1 <- points %>%
  mutate(nextLat = lead(position_lat),
         nextLng = lead(position_long),
         color = newcols.quantile
         )
gradient_map <- leaflet() %>% 
  addTiles()
points1 <- points1[points1$direction == 'Outbound',]
for (i in 1:nrow(points1)) {
  gradient_map <- addPolylines(map = gradient_map,
                               data = points1, 
                               lng = as.numeric(points1[i, c('position_long', 'nextLng')]), 
                               lat = as.numeric(points1[i, c('position_lat', 'nextLat')]), 
                               color = as.character(points1[i, c('color')])
  )
}

gradient_map
```
```{r}
newcols <- rev(cet_pal(min(dim(points)[1],256),name="l3"))
points <- points %>% mutate(quantile=ntile(heart_rate,256))
newcols.quantile <- newcols[points$quantile]
points1 <- points %>%
  mutate(nextLat = lead(position_lat),
         nextLng = lead(position_long),
         color = newcols.quantile
         )
gradient_map <- leaflet() %>% 
  addTiles()
points1 <- points1[points1$direction == 'Return',]
for (i in 1:nrow(points1)) {
  gradient_map <- addPolylines(map = gradient_map,
                               data = points1, 
                               lng = as.numeric(points1[i, c('position_long', 'nextLng')]), 
                               lat = as.numeric(points1[i, c('position_lat', 'nextLat')]), 
                               color = as.character(points1[i, c('color')])
  )
}

gradient_map
```

# Test with TrainingPeaks data

## WhiteLiam

Date: 16 April 2019, cycling

File: fit-test_cycling_2019-04-16_05-29-42.csv.fit

```{r}
data <- read.fit('fit-test_cycling_2019-04-16_05-29-42.csv.fit')
```


```{r}
# data$record$distance <- data$record$distance/100
 # #weird function to calculate correct altitude compared to tutorial
 # data$record$altitude <- unlist(lapply(data$record$altitude,function(x) x/(14.35-(x-3838)*6/2800)))
 names(data)
 head(data$record)
```


```{r}
attr(data$record,'units')
```

### Basic Plots

```{r}
pdata <- with(data$record, data.frame(alt = altitude, time = (timestamp-timestamp[1])/60))
 ggplot(pdata, aes(y=alt, x=time)) + geom_line() +
   ggtitle("Elevation vs Time") + xlab("time (minutes)") + ylab("elevation (m)")
```


```{r}
pdata <- with(data$record, data.frame(alt = altitude, time = (distance-distance[1])/1000))
 ggplot(pdata, aes(y=alt, x=time)) + geom_line() +
   ggtitle("Elevation vs Distance") + xlab("Distance (km)") + ylab("elevation (m)")
```

### Plot all variables Strava-like

```{r}
pdata <- with(data$record, data.frame(alt = altitude, time = (distance-distance[1])/1000,
                                      pow = power, hr=heart_rate, speed=speed, cad=cadence))
pdata <- pdata[-(1:10),]
pdata$grad <- with(data$record, 100 * diff(altitude,lag=10) / diff(distance,lag=10))

 ggplot(pdata, aes(y=alt, x=time)) +
   geom_line(aes(y=alt,colour="Altitude")) +
   geom_line(aes(y=pow*2,colour="Power"),size=0.05) +
   geom_line(aes(y=hr*2,colour="HR"),size=0.1) +
   geom_line(aes(y=speed*3.69*2,colour="Speed"),size=0.1) +
   geom_line(aes(y=cad*2,colour="Cadence"),size=0.05) +
   geom_line(aes(y=grad*10+min(pdata$alt),colour="Gradient"),size=0.2) +
   # geom_abline(intercept=min(pdata$alt), slope=0,color="orange",size=1,lty="longdash") +
   geom_segment(aes(x = 0, xend = max(pdata$time), y = min(pdata$alt), yend = min(pdata$alt)),color="orange",size=0.8) +
   scale_y_continuous(sec.axis = sec_axis(~./2, name = "HR (bpm) / Power (W) / Speed (kph) / Cadence(rpm)"),
                      breaks = seq(0, 1000, by = 200)) +
   scale_x_continuous(breaks= seq(0, 70, by = 10)) +
   ggtitle("Distance vs All") +
   xlab("Distance (km)") +
   ylab("Altitude (m)") +
   scale_colour_manual(values = c("black", "purple","orange","red","blue","green")) +
   labs(y = "Altitude (m)",
                x = "Distance (km)",
                colour = "") +
   theme_minimal() +
   theme(legend.position = "bottom")
   # theme(legend.position = c(0.9, 0.93))
```


```{r}
pdata <- data$record[-(1:10),c("heart_rate","timestamp")]
# compute average gradient, as %
 pdata$gradient <- with(data$record, 100 * diff(altitude,lag=10) / diff(distance,lag=10))
 pdata <- subset(pdata, complete.cases(pdata) & abs(gradient) < 7.5 & gradient != 0) # drop outliers
 ggplot(pdata, aes(x=gradient, y=heart_rate)) +
   geom_point(alpha=0.5) + geom_jitter() +
   stat_smooth(method="lm", formula=y ~ poly(x, 2)) +
   ggtitle("Heart rate vs gradient")
```

```{r}
 fit <- lm(heart_rate ~ poly(gradient, 2), data=pdata)
 summary(fit)
```


```{r}
pred <- data.frame(gradient = seq(-6,6,2))
pred$heart_rate <- predict(fit, pred)
t(pred)
```
 
```{r}
 # points <- subset(data$record, complete.cases(data$record))
points <- data$record
 points$time_min  <- with(points, timestamp - timestamp[1])/60 # minutes of riding

 # from diagram above, we turned around at the 90 minutes mark
 points[which(points$altitude == max(points$altitude)),]
 points$direction <- with(points, factor(ifelse(time_min < 90, 'Outbound', 'Return')))
```
```{r}
 library(cetcolor)
 ggplot(points,
        aes(y=position_lat, x=position_long, color=heart_rate)) +
   geom_path() + theme_dark() +
   facet_grid(direction~.) +
   xlab('Longitude') + ylab('Latitude') +
   ggtitle('Forerunner 230 GPS and Heart Rate Data') +
   scale_color_gradientn('Heart Rate (bpm)', colours = cet_pal(5, name="inferno")) +
   theme(plot.title = element_text(hjust=0.5, size=rel(2)),
         axis.title = element_text(size=rel(2)),
         legend.title = element_text(size=rel(1.5)))
   # ylim(c(-36.80,-36.65))
```

```{r}
 library(leaflet)
 leaflet(points[points$direction == 'Outbound',]) %>% addTiles() %>% addPolylines(~position_long,~position_lat)
 leaflet(points[points$direction == 'Return',]) %>% addTiles() %>% addPolylines(~position_long,~position_lat)
```

```{r}
newcols <- rev(cet_pal(min(dim(points)[1],256),name="l3"))
points <- points %>% mutate(quantile=ntile(heart_rate,256))
newcols.quantile <- newcols[points$quantile]
points1 <- points %>%
 mutate(nextLat = lead(position_lat),
        nextLng = lead(position_long),
        color = newcols.quantile
        )
gradient_map <- leaflet() %>%
 addTiles()
points1 <- points1[points1$direction == 'Outbound',]
for (i in 1:nrow(points1)) {
 gradient_map <- addPolylines(map = gradient_map,
                              data = points1,
                              lng = as.numeric(points1[i, c('position_long', 'nextLng')]),
                              lat = as.numeric(points1[i, c('position_lat', 'nextLat')]),
                              color = as.character(points1[i, c('color')])
 )
}

gradient_map
```


 
 
 
