  ---
title: "Training Peaks"
output:
  html_notebook:
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: true
    theme: cosmo
    df_print: paged
    highlight: tango
    code_folding: hide
    # fig_width: 12
    # fig_height: 12
# output:
#   epuRate::BAKER:
#     toc: TRUE
#     number_sections: FALSE
#     code_folding: "show"
---

<script>
$(document).ready(function() {
  $items = $('div#TOC li');
  $items.each(function(idx) {
    num_ul = $(this).parentsUntil('#TOC').length;
    $(this).css({'text-indent': num_ul * 10, 'padding-left': 0});
  });

});
</script>

***

# TrainingPeaks Data

## Preliminary tasks
### Load Libraries

Load libraries for reading fit data and create maps and plots.

```{r warning=FALSE, results='hide', echo=F}
# library(fit)
library(ggplot2)
library(leaflet)
library(dplyr)
library(cetcolor)
library(XML)
library(tidyr)
library(geosphere)
library(zoo)
library(RColorBrewer)
```


Read GPX file

```{r}
# load function
# https://rdrr.io/cran/plotKML/src/R/readGPX.R
source("readGPX.R")
```

```{r}
# read file
gpx <- readGPX("data/surfing.GPX")
# gpx <- readGPX("data/surfing_strava.gpx")
gpx <- gpx$tracks[[1]][[1]]
gpx
```

```{r}
# create variables with day and time
# ztime <- format(as.POSIXct(gpx$time[1],format="%Y-%m-%dT%H:%M:%S.000Z",tz="UTC"),tz="Australia/Melbourne")
# format(ztime)
### POLAR
gpx <- gpx %>%
  mutate(date=format(as.POSIXct(time,format="%Y-%m-%dT%H:%M:%S.000Z",tz="UTC"),tz="Australia/Melbourne"),
         day=format(as.POSIXct(date),"%y%m%d"),
         timeofday=format(as.POSIXct(date),"%H:%M:%S"),
         time=as.POSIXct(time,format="%Y-%m-%dT%H:%M:%S.000Z",tz="UTC")) %>%
  select(-date,-ele,-day)
### strava
# gpx <- gpx %>%
#   mutate(date=format(as.POSIXct(time,format="%Y-%m-%dT%H:%M:%SZ",tz="UTC"),tz="Australia/Melbourne"),
#          day=format(as.POSIXct(date),"%y%m%d"),
#          timeofday=format(as.POSIXct(date),"%H:%M:%S"),
#          time=as.POSIXct(time,format="%Y-%m-%dT%H:%M:%SZ",tz="UTC")) %>%
#   select(-date,-ele,-day)
# gpx
```

```{r}
# get distance and remove the steps with 0
gpx <- gpx %>% 
  mutate(laglon=lag(lon),
         laglat=lag(lat)) %>% 
  rowwise() %>%
  mutate(step_distance=distm(c(lon,lat),c(laglon,laglat),fun=distHaversine)) %>% 
  ungroup() %>% 
  mutate(step_distance=as.numeric(step_distance),
         cum_distance=round(cumsum(replace_na(step_distance,0)),2)) %>% 
  select(-laglon,-laglat) %>% 
  filter(step_distance != 0)

# get step time
gpx <- gpx %>% 
  mutate(step_s=as.numeric(time-lag(time)),
         step_s=case_when(
           is.na(step_s) ~ 1,
           step_s > 15 | lag(step_s,n=1) > 15 | lag(step_s,n=2) > 15 | lag(step_s,n=3) > 15  ~ rollmean(x=step_s,4,align="right",fill=NA),
           step_s > 10 | lag(step_s,n=1) > 10 | lag(step_s,n=2) > 10 ~ rollmean(x=step_s,3,align="right",fill=NA),
           step_s > 5 | lag(step_s,n=1) > 5 | lag(step_s,n=2) > 5 ~ rollmean(x=step_s,2,align="right",fill=NA),
           TRUE ~ round(step_s,2)
         ),
         cum_s=cumsum(replace_na(step_s,0)),
         cum_min=round(cum_s/60,2))
```


```{r}
# speed
gpx <- gpx %>% 
  mutate(speed.raw=(step_distance/1000)/(step_s/3600),
         speed=case_when(
           row_number() <= 10 ~ rollmean(x=speed.raw,10,align="left",fill=NA),
           TRUE ~ rollmean(x=speed.raw,10,align="right",fill=NA))
         )
plot(gpx$cum_distance,gpx$speed,type="l")
gpx
```

```{r}
wave <- gpx %>% slice(c(1:500)) %>% select(-speed.raw,-time,-step_distance)

min.speed <- 7
min.speedmax <- 13
max.speed <- 40 # for speed errors with GPS
min.length <- 6

wave <- gpx %>% 
  # slice(c(1600:1800)) %>% select(-speed.raw,-time,-step_distance) %>%
  mutate(wave.speed = ifelse(speed > min.speed,1,0)) %>% 
  mutate(wave.length = ifelse(wave.speed == 1 &
           (lag(wave.speed,n=5,default=0)+
           lag(wave.speed,n=4,default=0)+
           lag(wave.speed,n=3,default=0)+
           lag(wave.speed,n=2,default=0)+
           lag(wave.speed,n=1,default=0)+
           wave.speed+
           lead(wave.speed,n=1,default=0)+
           lead(wave.speed,n=2,default=0)+
           lead(wave.speed,n=3,default=0)+
           lead(wave.speed,n=4,default=0)+
           lead(wave.speed,n=5,default=0)) >= min.length,
           1,0)) %>%
  mutate(wave.n = ifelse(wave.speed != lag(wave.speed,default=0) & wave.speed == 1 & wave.length == 1,1,0)) %>%
  mutate(wave.n = ifelse(wave.speed == 1 & wave.length == 1,cumsum(replace_na(wave.n,0)),0)) %>%
  # filter(wave.speed !=0) %>% 
  group_by(wave.speed,wave.length,wave.n) %>%
  mutate(wave.speedcheck = case_when(
    any(wave.speed == 1 & wave.length==1 & speed > min.speedmax) & !any(speed.raw > max.speed & speed > min.speedmax) ~ 1,
    # any(wave.speed == 1 & wave.length==1 & speed > min.speedmax) ~ 1,
    TRUE ~ 0)) %>%
  ungroup() %>%
  mutate(wave.n = ifelse(wave.speed != lag(wave.speed,default=0) & wave.speed == 1 & wave.length == 1 & wave.speedcheck == 1,1,0)) %>%
  mutate(wave.n = ifelse(wave.speed == 1 & wave.length == 1 & wave.speedcheck == 1,cumsum(replace_na(wave.n,0)),0)) %>%
  data.frame()
wave
# wave <- wave %>% filter(wave.n != 0)
# wave
max(wave$wave.n)
# plot(wave$cum_distance,wave$speed,type="l")
```

```{r}
wave %>%filter(wave.n != 0)
```

################################
################################

*** 

```{r fig.width = 6}
# pdata <- with(data$record, data.frame(alt = altitude, time = (distance-distance[1])/1000,
#                                       pow = power, hr=heart_rate, speed=speed, cad=cadence))
# pdata <- pdata[-(1:10),]
# pdata$grad <- with(data$record, 100 * diff(altitude,lag=10) / diff(distance,lag=10))
mycols <- c("#000000",brewer.pal(n = 8, name = "Dark2"))
ggplot(wave, aes(y=speed, x=cum_distance, color=factor(wave.n))) +
  geom_line(color="black") +
  geom_line() +
  # breaks = seq(0, 1000, by = 200)) +
  # scale_x_continuous(breaks= seq(0, 70, by = 10)) +
  ggtitle("Distance vs All") +
  xlab("Distance (km)") +
  ylab("Altitude (m)") +
  scale_colour_manual(values=mycols) +
  # scale_colour_manual(values = c("black", "purple","orange","red","blue","green")) +
  labs(y = "Speed (km/h)",
       x = "Distance (m)",
       colour = "Wave") +
  theme_minimal() +
  # theme(legend.position = c(0.9, 0.93))
  theme(legend.position = "right")
```


### Activity Maps

Convert time to minutes and add direction (Outbound vs Return):
 
```{r}
points <- wave %>% select(lat,lon,speed,wave.n) %>%  filter(wave.n != 0)
# points$time_min  <- with(points, timestamp - timestamp[1])/60 # minutes of riding
```

```{r}
# library(leaflet)
leaflet(points) %>% addTiles() %>% addPolylines(~lon,~lat)
```

#### Map with HR

The darker the colour, the higher the HR is:

```{r}
# newcols <- rev(cet_pal(min(dim(points)[1],256),name="l3"))
points1 <- points %>%
  group_by(wave.n) %>% 
  mutate(nextLat = lead(lat),
         nextLng = lead(lon),
  ) %>% 
  ungroup() %>% 
  mutate(color=mycols[-1][as.factor(wave.n)])
# points1
gradient_map <- leaflet() %>%
 addTiles()
for (i in 1:nrow(points1)) {
 gradient_map <- addPolylines(map = gradient_map,
                              data = points1,
                              lng = as.numeric(points1[i, c('lon', 'nextLng')]),
                              lat = as.numeric(points1[i, c('lat', 'nextLat')]),
                              color = as.character(points1[i, c('color')])
 )
}
gradient_map
```